\begin{abstract}
The word \emph{data} is one of the most heavily used and least precisely
defined terms in computing. Standard references settle on an informal
characterization of data as ``raw facts,'' but none of them says what
structural properties a collection of facts actually needs before it can be
addressed, compared, or processed by a machine. This is not just a
philosophical loose end. It propagates directly into system design as schema
drift, type-confusion vulnerabilities, unsound equality checks, and
serialization defects, and each of these traces back to an unstated assumption
about what index set or value set a piece of data is supposed to inhabit. This
paper works through data, representation, and interpretation across three
registers. First, we ground the concept etymologically and lexically, tracing
the word \emph{datum} from its Latin root through to contemporary technical
usage, in order to expose a structural implication that the informal
definitions all share but never state outright. Second, we formalize data as a
\emph{function from an index set to a value set}, with explicit totality,
functionality, and codomain-consistency conditions, and we show how this
framing subsumes sequences, tuples, and multisets as special cases rather than
treating them as separate notions. Third, we introduce the
data--information--knowledge hierarchy as the interpretive scaffolding through
which machines, which operate at the data level and nothing more, are made to
produce outcomes that mean something to human beings, and we argue that
engineering this transition is really the central problem of data
representation. From the formal definition we derive four corollaries linking
specific violations of its conditions to recurring engineering failure modes,
namely schema drift, type confusion, unsound equality, and serialization
ambiguity, and we walk through a worked example for each. The paper closes by
separating data from representation, arguing that the choice of representation
is always contingent and sits outside the data itself, and that grasping this
distinction is a prerequisite for reasoning clearly about encoding, type
systems, and protocols.
\end{abstract}

\paragraph{Keywords.} data; representation; interpretation; formal definition;
data--information--knowledge hierarchy; type safety; schema drift;
serialization; systems design.
