\newpage
\section{A Formal Definition of Data}
\label{sec:formal-definition}

The informal definitions reviewed in Section~\ref{sec:what-is-data} are
consistent with each other, but none of them is formal. None says what
structure a collection of facts needs before it can be indexed, compared for
equality, or serialized. Mathematics gives us a sharper handle on the same
idea. In mathematics, data are most naturally thought of as a
\emph{function}: a mapping from some index set to some value set.

\subsection{Primitive Notions}

\begin{definition}[Index Set and Value Set]
\label{def:index-value}
An \emph{index set} $I$ is a set whose elements serve as positions or
addresses. A \emph{value set} $V$ is a set whose elements serve as the
possible contents at a position.
\end{definition}

Neither $I$ nor $V$ is constrained any further at this level of abstraction.
$I$ may be finite or infinite, ordered or unordered, and $V$ can contain
elements of any kind at all. What matters is just that both are fixed before a
collection of data gets constructed.

\begin{definition}[Datum]
\label{def:datum}
A \emph{datum} is an element $v \in V$, drawn from some value set $V$,
designated as the content held at a single index.
\end{definition}

\subsection{The Function-Theoretic Definition}

\begin{definition}[Collection of Data]
\label{def:collection}
A \emph{collection of data} $D$ over index set $I$ and value set $V$ is a
total function
\[
  D : I \to V
\]
assigning to every index $i \in I$ exactly one value $D(i) \in V$.
\end{definition}

For $D$ to be well-defined, three conditions need to hold.

\paragraph{Condition 1 (Totality).}
$D$ must be defined for every $i \in I$. A structure that leaves some indices
unassigned is a \emph{partial function}, and any operation that assumes
totality is unsound when applied to it without first checking the domain.

\paragraph{Condition 2 (Functionality).}
For every $i \in I$, $D(i)$ must denote a single, determinate value in $V$. A
mapping that assigns multiple candidate values to the same index, two fields
claiming the same byte offset under different schema versions, say, is not a
function, and therefore not a single well-defined collection of data.

\paragraph{Condition 3 (Codomain Consistency).}
Every value $D(i)$ must belong to the \emph{same} value set $V$ for all $i \in
I$, where $V$ is fixed in advance of constructing $D$.

\subsection{Equality}

The function-theoretic definition gives us a precise criterion for equality
that differs in an important way from mere representational identity.

\begin{definition}[Functional Equality]
\label{def:equality}
Two collections of data $D_1 : I_1 \to V_1$ and $D_2 : I_2 \to V_2$ are
\emph{equal}, written $D_1 = D_2$, if and only if $I_1 = I_2$, $V_1 = V_2$,
and $D_1(i) = D_2(i)$ for every $i \in I_1$.
\end{definition}

This is a stronger condition than byte-for-byte identity of an encoding. Two
structures can share an identical encoding while differing in $V$, exactly as
the five-byte example in Section~\ref{sec:what-is-data} shows, and
Definition~\ref{def:equality} correctly classifies them as distinct
collections of data. The reverse also holds: two collections can be equal
under Definition~\ref{def:equality} while differing in their encodings,
because the encoding plays no part in the definition at all.

\subsection{Relation to Standard Mathematical Structures}

The function-theoretic definition deliberately generalizes several structures
that get used informally as synonyms for ``data.''

\begin{description}
  \item[Sequence / Array.] The special case $I = \{0, 1, \dots, n-1\}$: a
        function from a finite, contiguous, totally ordered index set into
        $V$. This is the canonical definition of a sequence in the
        mathematical literature~\autocite{knuth1997}, and it is the formal
        starting point for arrays. Arrays are just data over the particular
        index set that integer addressing supplies.
  \item[Tuple.] The special case where $I$ is small and fixed, and $V$ is
        allowed to vary per index ($D(i) \in V_i$). Tuples relax Condition~3
        deliberately and openly, which is a structurally different move from
        the \emph{implicit} relaxation that produces type-confusion defects
        (Section~\ref{sec:implications}).
  \item[Multiset.] Corresponds to the case where $I$ is considered only up to
        a permutation. A multiset is a quotient of the function-theoretic
        structure, not some alternative to it: the underlying function still
        exists, and the multiset view simply ignores the ordering of $I$.
  \item[Relation / Table.] A relation with $k$ attributes over domains
        $V_1, \dots, V_k$ is, under this definition, a set of functions from
        $\{1,\dots,k\}$ into $V_1 \times \cdots \times V_k$. The relational
        model's primary key is a selection of indices into this function
        space.
\end{description}

The unifying point is that each of these structures is just a special case of
Definition~\ref{def:collection}, distinguished by the properties imposed on
$I$, $V$, or both. Moving between them is a matter of specifying or relaxing
constraints on those two sets, nothing more exotic than that.

\subsection{A Worked Example}

A weather station records temperature once per hour. Formally this is a
function $D : \mathbb{N} \to \mathbb{R}$, where the index set is the hour
number and the value set is temperature in degrees Celsius. The station's log
for one day is the restriction of $D$ to $I = \{0, 1, \dots, 23\}$, which is
an array. The highest temperature recorded is $\max_{i \in I} D(i)$, a
computation that is well-defined precisely because $I$ is finite and $D$ is
total on it. If a sensor fails and no reading gets recorded at hour 14, the
structure is no longer total on $I$, it has become a partial function, and any
operation that assumes totality, computing the daily average, for instance, is
unsound unless the missing value is handled explicitly.\\
This example shows why the three conditions matter in practice. Engineering
decisions about how to handle missing values, what type a field carries, and
whether two records actually count as the same record, all of these reduce to
questions about $I$, $V$, and the conditions in Definitions~\ref{def:collection}
and~\ref{def:equality}.
