\newpage
\section{Data, Information, and Knowledge}
\label{sec:data-information-knowledge}

The words \emph{data}, \emph{information}, and \emph{knowledge} get used
interchangeably in casual conversation, but they describe three distinct
things. The distinction matters in practice. Confusing them leads to systems
that produce numbers when people actually need answers. This section looks at
the relationship between the three concepts, then argues that computing
systems operate exclusively at the data level, and that the transition from
data to information is something engineers have to design deliberately rather
than get for free.

\subsection{The Hierarchy}

The relationship between data, information, and knowledge is standardly
described as a hierarchy~\autocite{ackoff1989,rowley2007}. This paper only
really depends on the first two transitions in that hierarchy.\\
Go back to the tree cross-section from Section~\ref{sec:what-is-data}. A raw
count of growth rings is data: fifty-three concentric lines, nothing more.
Data are \emph{syntactic}, they have structure but not yet meaning. Noticing
that fifty-three rings means the tree is roughly fifty-three years old, and
that the unusually narrow rings around year thirty line up with a known
regional drought, turns that count into \emph{information}: data organized and
placed in context, in keeping with Ratzan's definition~\autocite{ratzan2004}.
Information answers questions like ``what happened?'' and ``how much?'' It is
still factual and objective, but now it is interpretable by a human being.
Knowing how to read ring width as a record of annual rainfall, reliably enough
to apply the same method to a tree nobody has studied before, is
\emph{knowledge}: information understood thoroughly enough to be put to new
use.

\begin{example}
\label{ex:fever}
The progression from data to knowledge can be put concisely:
\begin{itemize}
  \item \textbf{Data:} \texttt{39.5}
  \item \textbf{Information:} Patient temperature is 39.5\textdegree C, recorded
        at 14:32 on 2026-06-28.
  \item \textbf{Knowledge:} This patient has a significant fever. Combined with
        readings from the past six hours showing a rising trend, the attending
        physician should be notified immediately.
\end{itemize}
The number \texttt{39.5} is just a datum drawn from $\mathbb{R}$. The
interpretive layer that attaches a unit, timestamp, and patient identity is
what constitutes information. The clinical judgment that combines this
information with prior cases and medical guidelines is what constitutes
knowledge.
\end{example}

\subsection{What Computing Systems Do}

Computers operate almost entirely at the level of \emph{data} as defined in
Section~\ref{sec:formal-definition}. A CPU does not understand information,
it manipulates bits. A memory array does not know what its contents mean, it
just stores and retrieves bytes at addresses. Shannon's foundational account
of communication is explicit about this: the engineering problem is the
reliable transmission of \emph{symbols}, regardless of what they
mean~\autocite{shannon1948}. Meaning is the problem of the sender and the
receiver, not the channel.\\
The transition from data to information is something software engineers build
into their systems on purpose, through type systems, schemas, protocols, and
documentation. When a programmer writes \texttt{int temperature = 39;}, what
they are really asserting is that this particular sequence of bits represents
a temperature, measured in some unit, at some precision. The meaning lives in
the programmer's head and in the documentation, not in the machine. The
machine just stores \texttt{0x00000027} and nothing else.\\
This is not a criticism of computers, it is just a description of how they
work, and understanding it is the first step toward seeing why encoding
schemes, type systems, and representation formats exist in the first place:
each one is an answer to the question of how to encode information, which has
meaning, as data, which does not, in a way that preserves the meaning
reliably.

\subsection{The Gap Easy Access Does Not Close}

Having access to information does not give you the understanding needed to
tell whether that information is relevant, reliable, or correctly applied. A
search engine can return information in milliseconds, but it cannot, on its
own, supply the judgment needed to evaluate it. The gap between having
information and having knowledge is not something retrieval speed closes.
This is an old observation, but it is still worth restating in a technical
context: systems that expose data as if it were information, or information
as if it were knowledge, end up misleading their users in predictable ways.
\newpage
\subsection{Relationship to the DIKW Hierarchy}

Ackoff's data--information--knowledge--wisdom hierarchy~\autocite{ackoff1989}
and its many descendants in the information-science
literature~\autocite{rowley2007,zins2007} consistently treat the data tier as
their foundation without ever defining it formally. The function-theoretic
definition in Section~\ref{sec:formal-definition} is meant as a precise
characterization of exactly that tier, one that is compatible with, and
presupposed by, every DIKW-style account. It does not compete with the
hierarchy. It fills the structural gap the hierarchy leaves open at the
bottom.

Similarly, Floridi's General Definition of Information
(GDI)~\autocite{floridi2010} characterizes information as data that are
well-formed, meaningful, and (in some variants) truthful, while treating data
themselves as a primitive that GDI deliberately leaves unspecified. The
function-theoretic definition supplies the structural detail that GDI leaves
open, without disturbing the relationship between data and information that
GDI describes.
