\section{Discussion}
\label{sec:discussion}

\subsection{Relationship to Existing Frameworks}

None of this is meant to compete with the GDI, with Zins' classification,
or with the DIKW hierarchy. Each of those frameworks leaves a gap open by
design, and the function-theoretic definition is built to sit inside that
gap, not replace what surrounds it. The contribution here is structural
precision at the data tier specifically---not a new theory of how data
becomes information, or how information becomes knowledge.

\subsection{Limitations}

Two cases push against the edges of what this definition assumes.

\paragraph{Streaming or unbounded data.}
When $I$ is not fixed in advance---an open-ended sensor stream is the
obvious example---Definition~\ref{def:collection}'s totality condition
cannot be checked against a complete, enumerable $I$, because there isn't
one yet. The definition still holds at any finite prefix or window of the
stream; the unbounded case just requires treating $I$ as a fixed,
well-ordered set such as $\mathbb{N}$, with totality reinterpreted as
``defined for every index produced so far'' rather than ``defined for
every index, full stop.''

\paragraph{Genuinely unstructured data.}
Free-text notes, before any tagging or field extraction has happened,
resist a useful index set on purpose. Treating such text as a function
from character position to character is technically valid---nothing in
the definition forbids it---but it throws away exactly the structure a
downstream NLP pipeline exists to recover. The definition is not wrong
here so much as premature: its value reappears once some later stage
imposes a more meaningful index set on the same text.

\subsection{Threats to Validity}

Every corollary above is illustrated with a constructed, representative
example, not pulled from an empirical defect-mining study. We treat the
corollaries as conceptually settled---each follows directly from
Definitions~\ref{def:collection} and~\ref{def:equality}, with nothing left
to argue about logically---but how large a share of real-world defects
they actually account for is an empirical question, and this paper does
not attempt to measure it.